Multi-stage programming with functors and monads: eliminating abstraction overhead from generic code

### 2007-07-12 SCP2008/paper/metamonads.tex
a few small changes. Far more changes tomorrow

\subsection{Abstract code and lifted domains}
We will be using MetaOCaml operations to build code expressions, values
of the type |('a,'v) code|. However, we wish to make our generators
abstract over the particular code representations. Therefore, in the
following Sections (in the rest of the paper?) rather than 
MetaOCaml brackets and escapes, we will be using code combinators
described below. These combinators are functions that produce and
combine values of the type of abstract code.

This abstraction lets us vary the output form of our generated LU
programs. We may use thunks rather than MetaOCaml |('a,'v) code|
values, for benchmarking purposes. We could use code combinators that
produce C or Fortran code.

@@

\jacques{some of these items for future work seem to be out of date, they need
to be reviewed}

We are looking into encapsulating staging
annotations into just a few functors, so that the rest of the code (in
particular, the |Gen| functor that puts it all together) should be
annotation-free and thus can be used as is in a one-stage environment
(pure OCaml) as well as in a multi-stage environment (generating
extensions). The one-stage code is a good baseline for benchmarks and
regression tests. Obtaining a generating extension from properly
modularized OCaml code (along the lines of our |Gen|) is an exciting
area of our future research.

@@

We plan to further investigate the connection between delimited
continuations and our implementations of code generators like
|ifM|.  As well, by using some additional syntactic sugar
(for |ifM|, |whileM|, etc.), the available notation should be
even more direct-style, and potentially clearer.
We also would like to extend our monad to a monad transformer.

>>>>

\subsection{Abstract code and lifted domains}
We will be using MetaOCaml operations to build code expressions, values
of the type |('a,'v) code|. However, we wish to make our generators
abstract over particular code representations. Therefore, in the
following Sections (in the rest of the paper?) instead of
MetaOCaml brackets and escapes we will be using code combinators
described below. These combinators are functions that produce and
combine values of the type of abstract code.

This encapsulation of staging annotations into a code combinators lets
us easily change the output form of our generated LU programs. We may
use single-stage thunks rather than MetaOCaml staged |('a,'v) code|
values, for benchmarking and regression tests purposes. We could use
code combinators producing C or Fortran code.

@@

@@

We plan to further investigate the connection between delimited
continuations and our implementations of code generators like
|ifM|. The ultimate (and plausible) goal is to write a numerical
algorithm in (almost) regular OCaml once, and being able to either
run it as a regular OCaml program, or turn into a code generating
aspect.
